---
title: "p_critical"
output: pdf_document
---

#################### PREFACE ##########################
#Code for PCA, NMDS, and PERMANOVA/ simper test 
#of raw pHMMer hits of citical carotenoid 
#biosynthesis enzymes: crtE, crtBa, crtBb, crtIa, crtIb, crtL, cruA, cruP
#cruF, and crtM. Both crtOb and crtZ have lots of 0's in
#their columns and can not be used when performing PCA. 
#These enzymes are create accessory carotenoid pigments 
#that could be important for photoprotection. 
#######################################################

```{r PCA}
#load packages and blastp raw hits data 
library(devtools)
library(ggbiplot)
library(readr)
p_critical <- read_csv("Nicole's Habitat Data Sets - p_critical.csv")
p_critical_matrix =data.matrix(p_critical[c(3:10)])
row.names(p_critical)=p_critical$Strain
pca_p_critical= prcomp(p_critical_matrix, center= TRUE, scale.= TRUE)

#analyze pca statistics to find proportion of variatiance explained by each gene 
summary(pca_p_critical)
str(pca_p_critical)

#make groups based on habitat
habitat= p_critical$Habitat

```

```{r PCA Plots, echo= FALSE}
#make the PCA plot with strains as points 
#png("PCA Plot of Carotenoid BLAST Hit .png", width=1000, height=1000)
ggbiplot(pca_p_critical,ellipse=FALSE,obs.scale = 3,var.scale = 1, groups=habitat) +
  scale_colour_manual(name="Habitat", values= c("cornflowerblue", "forest green", "darkorange1", "saddlebrown"))+
  ggtitle("PCA of pHMMer Critical Enzymes -N")+
  theme_minimal()+
  theme(legend.position = "bottom")

#make the PCA plot with strain names (looks too messy with all names...)
#png("PCA Plot of Carotenoid BLAST Hit .png", width=1000, height=1000)
ggbiplot(pca_p_critical,labels= rownames(p_critical),ellipse=FALSE,obs.scale = 3,var.scale = 1, groups=habitat) +
  scale_colour_manual(name="Habitat", values= c("cornflowerblue", "forest green", "darkorange1", "saddlebrown"))+
  ggtitle("PCA of pHMMer Critical Enzymes -N")+
  theme_minimal()+
  theme(legend.position = "bottom")

```

```{R NMDS}
#load packages 
library(vegan)
library(ggplot2)
library(readr)

#prep data
p_critical <- read_csv("Nicole's Habitat Data Sets - p_critical.csv")
p_critical_matrix =data.matrix(p_critical[c(3:10)])
nmds_p_critical_habitat= as.data.frame(p_critical[,1:2])
row.names(p_critical)=p_critical$Strain
#Do you need to transform your data? range should be 0-10 if not, use root function
range((p_critical_matrix)) # 0-12 should be 0-10 but lets see if the data looks fine 

#make a distance matrix
p_critical_matrix_ds=vegdist((p_critical_matrix), method = 'euclidian')

#Run metaMDS on distance matrix
p_critical_nms=metaMDS(p_critical_matrix_ds, distance="euclidian",k=2,trymax=500)
p_critical_nms
#stress=0.09

#assign colors to habitat
cols <- nmds_p_critical_habitat$Strain
cols[] <- c("cornflowerblue", "forest green", "darkorange1", "saddlebrown")
#Warning message:
#number of items to replace is not a multiple of replacement length

cols[nmds_p_critical_habitat$Habitat == c("aqutic")] <- "cornflowerblue"
cols[nmds_p_critical_habitat$Habitat == c("hydroterrestrial")] <- "forest green"
cols[nmds_critical_habitat$Habitat == c("soil")] <- "darkorange1"
cols[nmds_critical_habitat$Habitat == c("subaerial")] <- "saddlebrown"
cols[nmds_critical_habitat$Habitat == c("NA")] <- "grey"
pairs(character(nrow(nmds_p_critical_habitat$Strain, col=cols)))
              
habitatinfo= c(rep(p_critical$Habitat))
```

```{r NMDS Plots, echo= FALSE}
#NMDS Plot With Species Names 
ordiplot(p_critical_nms, type="n")
orditorp(p_critical_nms,display="sites",labels=p_critical$Strain, col=cols ,air=0.00025)
par(xpd=TRUE)
#legend(0.6,0.25,legend =unique(nmds_small_habitat$Strain), fill=cols, border= "black")
#change coordinates of the legend for each specific graph
#legend(-0.4,-0.2, "stress = 0.13") 
#change coordinates and stress value for each specific graph
title(main="NMDS of pHMMer critical Enzymes -N")
scores <-
  scores(p_critical_nms, display = "sites", "species")
cent <-
  aggregate(scores ~ habitatinfo, data = nmds_p_critical_habitat, FUN = "mean")
names(cent) [-1] <- colnames(scores)
      points(cent [,-1],
      pch = c( 8 , 8 , 8, 8),
      col = unique(cols),
      bg = c("black"),
      lwd = 3.0,
      cex = 2.0
)
                
#NMDS Plot with Points
ordiplot(p_critical_nms, choices= c(1,2))
points(p_critical_nms,display= "sites",pch=21, bg=cols) 
par(xpd=TRUE)
#legend(0.6,0.25,legend =habitat, fill=cols, border= "black")
#legend(-0.4,-0.2, "stress = 0.101") 
title(main="NMDS of pHMMer critical Enzymes -N")
scores <-
  scores(p_critical_nms, display = "sites", "species")
cent <-
  aggregate(scores ~ habitatinfo, data = nmds_p_critical_habitat, FUN = "mean")
names(cent) [-1] <- colnames(scores)
points(cent [,-1],
      pch = c( 8 , 8 , 8, 8),
      col = unique(cols),
      bg = c("black"),
      lwd = 3.0,
      cex = 2.0
)
```
                
```{R Stats}

#data prepartion 
rownames(p_critical_matrix) = p_critical$Strain
str(p_critical_matrix)
nmds_p_critical_habitat= as.data.frame(p_critical[,1:2])
str(nmds_p_critical_habitat)

#make a distance matrix
p_critical_matrix_ds=vegdist(p_critical_matrix, method = 'euclidian')

#Run metaMDS on distance matrix
p_critical_nms=metaMDS(p_critical_matrix_ds, distance="euclidian",k=2,trymax=500)
#stress=0.089

#Run PERMANOVA
pmv_p_critical = adonis(p_critical_matrix ~ habitat, data = nmds_p_critical_habitat, permutations = 999, method = 'bray')
pmv_p_critical
#PERMANOVA Results: Based off of the carotenoid genes being compared
#cyanobacterial strains differ statistically significant 
#between their habitats (p-value:).
#Habitats explian % of the variance in the copy numbers of critical 
#carotenoid biosynthesis enzymes between cyanobacterial species. 


#Visualise the permanova
densityplot(permustats(pmv_critical))

#Distance Based Dispersion test 
bd_p_critical = betadisper(p_critical_matrix_ds, p_critical$Habitat)
boxplot(bd_p_critical)
anova(bd_p_critical) 
#F-Test: pvalue 0.53
permutest(bd_p_critical)
#permutation test: pvalue 0.55

#both the F-Test and permutation test show that 
#the avergae distance to the median for the 
#habitat data is not statistically significant. 

#simper test: determine which genes are responsible for the difference between 2 compared habiatat groups 
simp_p_critical= simper(p_critical_matrix, group= p_critical$Habitat)
summary(simp_p_critical)
#Results of simper test
```
knitr::opts_chunk$set(echo = TRUE)
```