---
title: "Carotenoid Gene Trees pHMMer"
output: pdf_document
---

```{r attempt #1 from youtube video}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggtree)
library(reshape2)
library(ggstance)
library(treeio)
library(tidytree)
library(ape)
setwd("/bigdata/stajichlab/shared/projects/Cyanobacteria/Cyanobacteria_pigments/genetree_pHMMer")


#load tree data 
#NOTE: must convert nexus files to newick, can do this in figtree 
mytree_crtBa <- 'crtBa.newick'
tree_crtBa <- read.tree(mytree_crtBa)

#import group metadata
library(readr)
crtBagroups <- read.table("Nicole's Habitat Data Sets - crtBa_groups.tsv", sep = '\t', 
                         col.names = c('Tip_label', 'Habitat'),
                         header = FALSE)
crtBagroups

#convert group metadata to R data frame and combine with tree data 
gdata_crtBa = as.data.frame(crtBagroups)
crtBa_tree1 = full_join(as_tibble(tree_crtBa), gdata_crtBa, by = c('label'='Habitat'))
crtBa_tree2 = as.treedata(crtBa_tree1)

#Warning message:`mutate_()` was deprecated in dplyr 0.7.0. Please use `mutate()` instead. See vignette('programming') for more help --> make sure to have Ape package! 

crtBa_color <- ggtree(crtBa_tree2) + geom_tiplab(aes(color =Habitat)) + xlim(-5,5)
```

``````{r attempt #2}
#load packages and set wd 
library(ggtree)
set.seed(2017-02-16)
setwd("/bigdata/stajichlab/shared/projects/Cyanobacteria/Cyanobacteria_pigments/genetree_pHMMer")

#load tree data 
#NOTE: must convert nexus files to newick, can do this in figtree 
mytree_crtBa <- 'crtBa.newick'
tree_crtBa <- read.tree(mytree_crtBa)
ggtree(tree_crtBa) + geom_tiplab(align = TRUE, linesize = 0.25)

#import group metadata
#library(readr)
groups_all <- read_csv("Cyanobacteria and Carotenoids Metadata - groups_all.csv")
groups_all
crtBa_tiplabels <- TipLabels(mytree_crtBa)